Loop Invariant Code Motion in PyPy's Tracing JIT

### 2011-06-16 talk/iwtc11/paper.tex
make clear that we don't claim novelty

In this paper we want to address this problem by proposing a simple scheme that
makes it possible to turn simple optimizations using one forward pass into
optimizations that can do loop invariant code motion and similar loop-aware
improvements. Using this scheme one does not need to change the underlying
optimization much to get these advantages.

@@

Loop peeling is achieved by appending an copy of the traced iteration at
the end of itselfe. See Figure~\ref{fig:overview} for an illustration.
The first part (called \emph{preamble}) finishes with the jump the the second part
(called the \emph{peeled loop}). The second part end with the jump to itself. This way
the preamble will be executed only once while the peeled loop will
be used for every further iteration. New variable names have to be
introduced in the entire copied trace in order to maintian the SSA-property.
When applying the following optimizations to this two-iteration trace
some care has to taken as to how the arguments of the two
\lstinline{jump} operations and the input arguments of the peeled loop are
treated. It has to be ensured that the peeled loop stays a proper
trace in the sense that the operations within it only operates on
variables that are either among its input arguments 
or produced within the peeled loop. To ensure this we need
to introduce a bit of formalism.

>>>>

In this paper we want to address this problem by proposing a simple scheme that
makes it possible to turn optimizations using one forward pass into
optimizations that can do loop invariant code motion and similar loop-aware
improvements. Using this scheme one does not need to change the underlying
optimization much to get these advantages.

The resulting optimizations one gets using this scheme are in no way novel, most
of them are well-known loop optimizations. However, the way to implement them is
a lot simpler than directly implementing loop-aware optimizations.

@@

Loop peeling is achieved by appending an copy of the traced iteration at
the end of itself. See Figure~\ref{fig:overview} for an illustration.
The first part (called \emph{preamble}) finishes with the jump the the second part
(called the \emph{peeled loop}). The second part end with the jump to itself. This way
the preamble will be executed only once while the peeled loop will
be used for every further iteration. New variable names have to be
introduced in the entire copied trace in order to maintian the SSA-property.
When applying the following optimizations to this two-iteration trace
some care has to taken as to how the arguments of the two
\lstinline{jump} operations and the input arguments of the peeled loop are
treated. It has to be ensured that the peeled loop stays a proper
trace in the sense that the operations within it only operates on
variables that are either among its input arguments 
or produced within the peeled loop. To ensure this we need
to introduce a bit of formalism.
